fGFPCA dynamic prediction (In-sample)

### 2022-12-22 f-GFPCA.Rmd
start on out-of-sample pred

---
title: "fGFPCA dynamic prediction (In-sample)"
output: html_document
date: "2022-11-03"
---

@@

## Toy simulation

### Simulation set-up

@@

### Local GLMMs

@@

### FPCA

@@

### In-sample prediction

@@

@@

```{r}
source(here("code/DynPred.R"))
```

```{r}
# function to clean results
# make predictions with partial track up to 20, 40, 60
clean_pred <- function(subj = 1){
  pred20 <- data.frame(bin = mid, 
                       eta_i = in_samp_dyn_pred(fpca_fit, 20,
                                                df_pred_unique, subj),
                       type = "up to 195") %>% 
    filter(bin>195)

  pred40 <- data.frame(bin = mid, 
                      eta_i = in_samp_dyn_pred(fpca_fit, 40,
                                               df_pred_unique, subj),
                       type = "up to 395") %>%
    filter(bin>395)

  pred60 <- data.frame(bin = mid, 
                       eta_i = in_samp_dyn_pred(fpca_fit, 60,
                                                df_pred_unique, subj),
                        type = "up to 595") %>%
    filter(bin>595)

  df_pred <- df_pred_latent %>% filter(id==subj & sind_inx==bin) %>%
    select(bin, eta_i) %>% mutate(type="true") %>%
    add_row(pred20) %>%
    add_row(pred40) %>% 
    add_row(pred60)

@@

```{r}
# Try on subjects 1-4
ggarrange(
clean_pred(1) %>%
  ggplot()+
  geom_line(aes(x=bin, y = eta_i, col=type))+
  labs(title = "Subject 1"),

clean_pred(2) %>%
  ggplot()+
  geom_line(aes(x=bin, y = eta_i, col=type))+
  labs(title = "Subject 2"),

clean_pred(3) %>%
  ggplot()+
  geom_line(aes(x=bin, y = eta_i, col=type))+
  labs(title = "Subject 3"),

clean_pred(4) %>%
  ggplot()+
  geom_line(aes(x=bin, y = eta_i, col=type))+
  labs(title = "Subject 4"), nrow=2, ncol=2, common.legend = T)

>>>>

---
title: "fGFPCA dynamic prediction (In-sample)"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
date: "2022-11-03"
---

@@

# Toy simulation

## Simulation set-up

@@

## Local GLMMs

@@

## FPCA

@@

## In-sample prediction

@@

Let

$$\boldsymbol{H} = \boldsymbol{\hat{\Gamma}\Phi ^T_m}(\boldsymbol{\Phi_m\hat{\Gamma}\Phi^T_m}+\hat{\sigma_{\epsilon}}^2\boldsymbol{I_m})^{-1}$$

The variance of $\hat{\boldsymbol{\eta}}$ can be estimated by:

$$Var(\hat{\boldsymbol{\eta}}-\boldsymbol{\eta})=\boldsymbol{\Phi}(\hat{\boldsymbol{\Gamma}}-\boldsymbol{H}\boldsymbol{\Phi_m}\hat{\boldsymbol{\Gamma}})\boldsymbol{\Phi}^T+\hat{\sigma_{\epsilon}}^2$$

It should be a $S \times S$ matrix; diagonals are variance at a specific time and off-diagonals are covariance between different times on binned grid.

@@

```{r}
source(here("code/DynPred.R"))

#in_samp_dyn_pred(fpca_fit, 20, df_pred_unique, 1)

```{r}
# function to clean results
# make predictions with partial track up to 20, 40, 60
clean_pred <- function(subj = 1){

  pred20 <- in_samp_dyn_pred(fpca_fit, 20, df_pred_unique, subj)
  df_pred20 <- data.frame(bin = mid, 
                          eta_i = pred20$eta_pred,
                          eta_i_std=pred20$eta_std_var,
                          type = "up to 195") 
  df_pred20$eta_i[df_pred20$bin<=195] <- NA
  df_pred20$eta_i_std[df_pred20$bin<=195] <- NA

  pred40 <- in_samp_dyn_pred(fpca_fit, 40, df_pred_unique, subj)
  df_pred40 <- data.frame(bin = mid, 
                          eta_i = pred40$eta_pred,
                          eta_i_std=pred40$eta_std_var,
                          type = "up to 395") 
  df_pred40$eta_i[df_pred20$bin<=395] <- NA
  df_pred40$eta_i_std[df_pred20$bin<=395] <- NA

  pred60 <- in_samp_dyn_pred(fpca_fit, 60, df_pred_unique, subj)
  df_pred60 <- data.frame(bin = mid, 
                          eta_i = pred60$eta_pred,
                          eta_i_std=pred60$eta_std_var,
                          type = "up to 595")
  df_pred60$eta_i[df_pred20$bin<=595] <- NA
  df_pred60$eta_i_std[df_pred20$bin<=595] <- NA

  df_pred <- df_pred20 %>%
    add_row(df_pred40) %>% 
    add_row(df_pred60) %>%
    mutate(eta_i_lower = eta_i-eta_i_std, 
           eta_i_upper = eta_i+eta_i_std) %>%
    left_join(df_pred_latent %>% 
                 filter(id==subj & sind_inx==bin) %>%
                 select(bin, eta_i) %>% 
                 rename(true_eta = eta_i), by = "bin")

@@

```{r, fig.width=12, fig.height=4}
# Try on subjects 1-4
clean_pred(1) %>%
  ggplot()+
  geom_line(aes(x=bin, y = true_eta), col = "red")+
  geom_line(aes(x=bin, y = eta_i), na.rm = T)+
  geom_line(aes(x=bin, y=eta_i_lower), linetype = "dashed", na.rm=T)+
  geom_line(aes(x=bin, y = eta_i_upper), linetype = "dashed", na.rm=T)+
  labs(title = "Subject 1", x = "bin", y="")+
  facet_wrap(~type)

clean_pred(2) %>%
  ggplot()+
  geom_line(aes(x=bin, y = true_eta), col = "red")+
  geom_line(aes(x=bin, y = eta_i), na.rm=T)+
  geom_line(aes(x=bin, y=eta_i_lower), linetype = "dashed", na.rm=T)+
  geom_line(aes(x=bin, y = eta_i_upper), linetype = "dashed", na.rm=T)+
  labs(title = "Subject 2", x = "bin", y="")+
  facet_wrap(~type)

clean_pred(3) %>%
  ggplot()+
  geom_line(aes(x=bin, y=true_eta), col = "red")+
  geom_line(aes(x=bin, y = eta_i), na.rm = T)+
  geom_line(aes(x=bin, y=eta_i_lower), linetype = "dashed", na.rm=T)+
  geom_line(aes(x=bin, y = eta_i_upper), linetype = "dashed", na.rm=T)+
  labs(title = "Subject 3", x = "bin", y="")+
  facet_wrap(~type)

clean_pred(4) %>%
  ggplot()+
  geom_line(aes(x=bin, y=true_eta), col = "red")+
  geom_line(aes(x=bin, y = eta_i), na.rm = T)+
  geom_line(aes(x=bin, y=eta_i_lower), linetype = "dashed", na.rm=T)+
  geom_line(aes(x=bin, y = eta_i_upper), linetype = "dashed", na.rm=T)+
  labs(title = "Subject 4", x = "bin", y="")+
  facet_wrap(~type)
```

## Out of sample prediction

Now let's say if we have a new subject with observations up to $t_m$ and $t_m$ is in mth bin:

1. Local GLMMs: we will be able to estimate random effect of this subject up to mth bin, thus estimate the latent function up to the mth bin ($\hat{\boldsymbol{\eta}}_m$).

2. fPCA: We will have an incomplete function for this new observation. I think fpca.face can handle that.

3. Use the save formulas as in-sample prediction
